---
title: "Dataframe Summary"
output: "pdf_document"
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)

format_perc <- function(x) {
    sprintf("%.1f", x)
}

format_exp <- function(x) {
    sprintf("%.1e", x)
}

read_tsv <- function(path) {
    read.csv(path, sep = "\t", na.strings = c(".")) %>%
        as_tibble()
}

make_stats_table <- function(df) {
    gather(df) %>%
        group_by(key) %>%
        summarize(min = min(value, na.rm = TRUE),
                  max = max(value, na.rm = TRUE),
                  med = median(value, na.rm = TRUE),
                  mean = mean(value, na.rm = TRUE),
                  stdev = mean(value, na.rm = TRUE),
                  range = max - min,
                  n_NA = sum(is.na(value)),
                  perc_NA = 100 * n_NA / n()) %>%
        rename(feature = key) %>%
        mutate(perc_NA = sprintf("%.1f", perc_NA)) %>%
        mutate(across(c(min, max, med, mean, stdev, range), format_exp))
}

make_feature_distribution <- function(x) {
    ggplot() +
        aes(x) +
        geom_density() +
        xlab(NULL) +
        ylab(NULL)
}

make_cor_plot <- function(df) {
    cor(df) %>%
        as_tibble(rownames = "var1") %>%
        gather(-var1, key = var2, value = coeff) %>%
        ggplot(aes(var1, var2, fill = coeff)) + geom_tile() +
        scale_fill_gradient2(low = "red",
                             mid = "black",
                             high = "green",
                             limits = c(-1, 1)) +
        xlab(NULL) +
        ylab(NULL) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1),
              legend.position = "bottom")
}

make_colocalization_plot <- function(df) {
    mutate(df, across(everything(), ~ if_else(is.na(.x), 0, 1))) %>%
        make_cor_plot()
}

df <- read_tsv(snakemake@input[[1]])

df_x <- df %>%
    select(-label) %>%
    select(-CHROM, -POS, -POS.length.REF., -FILTER, -GT, -GQ)

df_y <- df$label

```

# Dataframe Overview

```{r, echo=FALSE}
N <- length(df_y)
P <- length(names(df_x))
N_TP <- length(keep(df_y, ~ .x == "tp"))
perc_TP <- sprintf("%.2f", 100 * N_TP / N)
```

* N Rows: `r N`
* N Features: `r P`
* N True Positives: `r N_TP`
* Perc True Postives: `r perc_TP`

# Feature Summary table

```{r, results="asis", echo=FALSE}

df_x %>%
    make_stats_table() %>%
    knitr::kable()

```

# Colocalization Matrix

```{r, echo=FALSE}

no_missing <- df_x %>%
    gather() %>%
    group_by(key) %>%
    summarize(value = any(is.na(value))) %>%
    filter(value == FALSE) %>%
    pull(key)

```

All missing values are converted to 0, and all present values are converted to
1, and the resulting correlation matrix is computed.

Interpretation:
* coeff = 1 means perfect colocalization (each value either missing or present)
* coeff = -1 means mutually exclusive

The following features had no missing values: `r paste(no_missing, sep=", ")`

```{r, echo=FALSE, fig.width = 7, fig.height = 7}

df_x %>%
    select(-all_of(no_missing)) %>%
    make_colocalization_plot()

```

# Feature distributions

Plots of all features with missing values removed.

```{r, echo=FALSE, results="asis", fig.width=4, fig.height=4}

print_plot <- function(x, name) {
    cat(sprintf("## %s\n", name))

    .x <- discard(x, is.na)

    if (max(.x) - min(.x) == 0) {
        cat(sprintf("Feature has one value: %.1f", max(.x)))
    } else {
        print(make_feature_distribution(.x))
    }

    cat("\n\n")
}

iwalk(df_x, ~ print_plot(.x, .y))

```
