---
title: "Dataframe Summary"
output: "html_document"
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(infotheo)

# from blablabla import nukes
`:=` = rlang::`:=`
`!!` = rlang::`!!`

## gross...
source("workflow/scripts/rmarkdown/common/colocalization.r")
source("workflow/scripts/rmarkdown/common/plots.r")

format_perc <- function(x) {
    sprintf("%.4f", x)
}

format_exp <- function(x) {
    sprintf("%.1e", x)
}

make_stats_table <- function(df) {
    N <- nrow(df)
    gather(df, factor_key = TRUE) %>%
        group_by(key) %>%
        summarize(n_present = sum(!is.na(value)),
                  perc_present = 100 * n_present / N,
                  # prevent NULL error for zero length vectors in min/max
                  min = ifelse(n_present == 0, NA, min(value, na.rm = TRUE)),
                  max = ifelse(n_present == 0, NA, max(value, na.rm = TRUE)),
                  med = median(value, na.rm = TRUE),
                  mean = mean(value, na.rm = TRUE),
                  stdev = sd(value, na.rm = TRUE),
                  range = max - min) %>%
        rename(feature = key) %>%
        mutate(perc_present = format_perc(perc_present)) %>%
        mutate(across(c(min, max, med, mean, stdev, range), format_exp)) %>%
        arrange(desc(as.numeric(perc_present))) %>%
        knitr::kable(align = "r")
}

make_feature_distribution <- function(x, labels) {
    infer_transform(x) %>%
        mutate(label = labels) %>%
        gather(-label, key = key, value = value) %>%
        ggplot() +
        aes(value, color = label) +
        geom_density() +
        xlab(NULL) +
        ylab(NULL) +
        facet_wrap(~key, scales = "free")
}

label_summary_table <- function(y) {
    tibble(label = y) %>%
        group_by(label) %>%
        summarize(n = n(),
                  proportion = format_perc(n / N)) %>%
        knitr::kable()
}

fmt_vcf_col <- function(fconf, which) {
    .v <- fconf[["vcf"]]
    sprintf("%s_%s", .v[["prefix"]], .v[["columns"]][[which]])
}

path <- snakemake@input[[1]]

fconf <- snakemake@config[["features"]]
label_col <- fconf[["label"]]
idx <- fconf[["index"]]

df_x <- readr::read_tsv(path,
                        col_types = cols(
                            !!label_col := "-",
                            !!idx[["chr"]] := "-",
                            !!idx[["start"]] := "-",
                            !!idx[["end"]] := "-",
                            !!fmt_vcf_col(fconf, "filter") := "-",
                            !!fmt_vcf_col(fconf, "gt") := "-",
                            !!fmt_vcf_col(fconf, "gq") := "-",
                            .default = "d"))

df_y <- readr::read_tsv(path, col_types = cols(
                                  !!label_col := "c",
                                  .default = "-")) %>%
    pull(!!label_col)

y_labels <- unique(df_y)

features <- names(df_x)

```

# Dataframe Overview

```{r, echo=FALSE}
N <- length(df_y)
P <- length(names(df_x))
```

* N Rows: `r N`
* N Features: `r P`

## Label fractions

```{r, echo=FALSE, results = "asis"}
label_summary_table(df_y)
```

# Feature Summary Tables

Ranked by descending percent presence

## Combined Labels

```{r, results = "asis", echo = FALSE}

df_x %>%
    make_stats_table()

```

```{r, results = "asis", echo = FALSE}

print_label_tbl <- function(label) {
    cat(sprintf("## %s Label\n\n", label))

    df_x %>%
        filter(df_y == label) %>%
        make_stats_table() %>%
        print()

    cat("\n\n")
}

walk(as.list(y_labels), print_label_tbl)

```

# Feature ranking

For each feature, compute the mutual information between the feature and the
label, then normalize by the entropy of the label. The resulting value can be
interpreted as the amount of information of the label that is also in the
feature (although the feature may have additional information outside the label
since many features are not binary).

Do this for all features when either filtering out missing values or replacing
missing values with 0.

```{r include = FALSE, message = FALSE }

information <- function(df, var1, var2) {
    .df <- df %>%
        select(all_of(c(var1, var2))) %>%
        drop_na()
    n <- nrow(.df)
    if (n > 0) {
        nbreaks <- sqrt(n) %>% ifelse(. > 2, ., 2)
        mi <- .df %>%
            ## the discretize function doesn't seem to work the way I want, so
            ## just use 'cut' since I know what that does
            ## discretize() %>%
            mutate(across(everything(),
                          ~ cut(.x, breaks = nbreaks, labels = FALSE) %>%
                              as.vector())) %>%
            mutinformation()
        H1 <- mi[1, 1]
        H2 <- mi[2, 2]
        I <- mi[1, 2]
    } else {
        H1 <- NA
        H2 <- NA
        I <- NA
    }
    list(H1 = H1,
         H2 = H2,
         ## mutual information
         I = I,
         ## mutual information normalized to joint entropy
         Inorm = I / (H1 + H2 - I),
         ## mutual information normalized to the first feature
         I_H1 = I / H1,
         ## variation of information (if a metric is needed)
         VI = H1 + H2 - 2 * I)
}

info_df <- function(features, df_info) {
    features %>%
        as.list() %>%
        map(~ information(df_info, "label", .x)) %>%
        tibble(i = ., param = features) %>%
        unnest_wider(i) %>%
        drop_na()
}

info_plot <- function(df) {
    ggplot(df, aes(reorder(param, desc(I_H1)), I_H1)) +
        geom_col() +
        xlab(NULL) +
        ylab("Mutual Inf. (Normalized to Label)") +
        theme(axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90))
}

print_info_plot <- function(df_info, features) {
    info_df(features, df_info) %>%
        info_plot() %>%
        print()
}

df_info_na <- df_x %>%
    mutate(label = as.integer(df_y == "tp"))

df_info_filled <- df_info_na %>%
    mutate(across(everything(), ~ if_else(is.na(.x), 0.0, as.double(.x))))

```

## Replace missing with 0

```{r, echo = FALSE, results = "asis", fig.width = 8, fig.height = 5}

print_info_plot(df_info_filled, features)

```

## Filter missing

```{r, echo = FALSE, results = "asis", fig.width = 8, fig.height = 5}

print_info_plot(df_info_na, features)

```

# Colocalization Matrices

Convert all features into binary values (TRUE is "not missing") then calculate
the "assymmetric jaccard index" for all feature pairs.

Formula for assymetric jaccard index for sets $A$ and $B$:

$$j = \frac{\lvert A \cap B \rvert}{\lvert A \rvert}$$

Plot interpretation: Given feature $A$ on the x axis and feature $B$ on the
y axis, $B$ overlaps with $A$ according to the value of their intersecting
cell.

```{r, include = FALSE}

ajaccard_combinations <- function(df_bool, comb_df) {
}

print_coloc <- function(df_bool, comb_df) {
    mutate(comb_df, asymm_jaccard = ajaccard(df_bool, var.x, var.y)) %>%
        make_xy_tile_plot("var.x",
                          "var.y",
                          "asymm_jaccard",
                          "starting set",
                          "overlapping set") %>%
        print()

    cat("\n\n")
}

combinations <- df_x %>%
    names() %>%
    cross_tibble()

df_x_bool <- to_binary(df_x)

```

```{r, echo=F, results = "asis", fig.width = 10, fig.height = 10}

print_label_coloc <- function(label) {
    cat(sprintf("## %s only\n\n", label))

    df_x_bool <- filter(df_x, df_y == label) %>%
        to_binary()
    print_coloc(df_x_bool, combinations)

    cat("\n\n")
}

cat("## All labels\n\n")

print_coloc(df_x_bool, combinations)

walk(as.list(y_labels), print_label_coloc)

```

# Correlation matrices

Find all features where the jaccard index (the real one, not our weird made-up
assymetric version) is 1 (perfect overlap). Partition those features by those
that are in combination with each other, and then compute the correlation matrix
on these subsets after filtering missing values.

```{r, include = FALSE}

## ASSUME these will be the same for TP/FP/both
perfect_overlaps <- df_x_bool %>%
    perfect_overlapping_sets(combinations, "var.x", "var.y")

print_subset_cor_plot <- function(df, subset) {
    df %>%
        select(all_of(subset)) %>%
        drop_na() %>%
        make_cor_plot() %>%
        print()
}

print_cor_plots <- function(df, subsets) {
    cat(sprintf("number of rows: %s\n\n", nrow(df)))

    walk(as.list(subsets), ~ print_subset_cor_plot(df, .x))

    cat("\n\n")
}

```

```{r, echo=FALSE, results = "asis", fig.width = 7, fig.height = 7}

print_label_cor <- function(label) {
    cat(sprintf("## %s only\n\n", label))
    filter(df_x, df_y == label) %>% print_cor_plots(perfect_overlaps)
    cat("\n\n")
}

cat("## All labels\n\n")

print_cor_plots(df_x, perfect_overlaps)

walk(as.list(y_labels), print_label_cor)

```

# Feature distributions

Plots of all features with missing values removed.

The following transforms were used:

- log10: `log(x, 10)`
- arsinh: `arsinh(x)`

Transforms were used depending on the domain of the input vector:

- 0 to 1 (inclusive): no transform
- greater than 0: log10
- all reals: arsinh

```{r, echo=FALSE, results="asis", fig.width=8, fig.height=4}

print_plot <- function(x, name) {
    cat(sprintf("## %s\n\n", name))


    .df <- tibble(x = x, y = df_y) %>%
        filter(!is.na(x))

    label_summary_table(.df$y) %>% print()

    cat("\n\n")

    .x <- .df$x
    .y <- .df$y

    if (length(.x) == 0) {
        cat("Feature has no values")
    } else if (max(.x) - min(.x) == 0) {
        cat(sprintf("Feature has one value: %.1f", max(.x)))
    } else {
        print(make_feature_distribution(.x, .y))
    }

    cat("\n\n")
}

iwalk(df_x, ~ print_plot(.x, .y))

```
