---
title: "Model Summary"
output: "html_document"
---

```{r, include=FALSE, message=FALSE}
library(tidyverse)
library(ggpubr)

source("workflow/scripts/rmarkdown/common/plots.r")

to_tibble <- function(lst) {
    do.call(tibble, lst)
}

read_model <- function(path) {
    jsonlite::read_json(path, simplifyVector = TRUE, simplifyDataFrame = FALSE)
}

read_predictions <- function(path) {
    readr::read_tsv(path, col_types = cols(.default = "d"))
}

lookup_input_path <- function(mapping, k) {
    pluck(mapping, as.character(as.integer(k)))
}

to_univariate <- function(model) {
    model$univariate %>%
        map(~ list(meta = .x[c("name", "vartype")], df = to_tibble(.x[["df"]])))
}

to_bivariate <- function(model) {
    model$bivariate %>%
        map(~ list(left = .x$left, right = .x$right, df = to_tibble(.x$df)))
}

run_features <- snakemake@params[["features"]]
error_labels <- snakemake@params[["error_labels"]]

mod <- read_model(snakemake@input[["model"]])
test_pred <- read_predictions(snakemake@input[["predictions"]])
train_pred <- read_predictions(snakemake@input[["train_predictions"]])

train_x <- readr::read_tsv(snakemake@input[["train_x"]],
                           col_types = cols(chrom = "-",
                                            chromStart = "-",
                                            chromEnd = "-",
                                            variant_index = "-",
                                            .default = "d"))
train_y <- readr::read_tsv(snakemake@input[["train_y"]],
                           col_types = cols(.default = "d"))

threshold <- train_pred %>% pull(label) %>% mean()

alltrain <- bind_cols(train_x, train_pred) %>%
    mutate(pred = prob > threshold)

VCF_input_name <- "VCF_input"

global_df <- to_tibble(mod$global_scores)

univariate <- to_univariate(mod)
bivariate <- to_bivariate(mod)

```

# Model Overview

* Variant Type: `r snakemake@wildcards$filter_key`
* EBM Run Key: `r snakemake@wildcards$run_key`
* Error Labels: `r str_c(error_labels, collapse = ", ")`

# Model Statistics

## Input file keys:

```{r, echo = F, results = "asis"}

cat("\n")

```

## Overview

* N: `r nrow(test_pred)`
* Perc. Pos: `r sum(test_pred$label)/nrow(test_pred)`

## Prediction Distribution

```{r, echo = FALSE, results = "asis", fig.width = 7, fig.height = 4}
ggplot(test_pred, aes(prob, color = factor(label))) +
    geom_density() +
    xlab("probability") +
    scale_color_discrete(name = "label")
```

## Calibration

```{r, echo = FALSE, results = "asis", fig.width = 4, fig.height = 4}

nbins <- 10

test_pred %>%
    mutate(bin = cut(prob, nbins, labels = FALSE) / nbins) %>%
    group_by(bin) %>%
    summarize(mean_pred = mean(prob), frac_pos = mean(label)) %>%
    ggplot(aes(mean_pred, frac_pos)) +
    geom_point() +
    geom_line() +
    geom_abline(linetype = "dotted", color = "red") +
    xlim(0, 1) +
    ylim(0, 1)

```

## ROC Curve

```{r, include=FALSE}

AllP <- sum(test_pred$label)
AllN <- nrow(test_pred) - AllP

roc <- test_pred %>%
    arrange(prob) %>%
    mutate(thresh_FN = cumsum(label),
           thresh_TN = row_number() - thresh_FN,
           thresh_TP = AllP - thresh_FN,
           thresh_FP = AllN - thresh_TN,
           TPR = thresh_TP / AllP,
           TNR = thresh_TN / AllN,
           FPR = 1 - TNR,
           precision = thresh_TP / (thresh_TP + thresh_FP))

```

```{r, echo = FALSE, results = "asis", fig.width = 4, fig.height = 4}
roc %>%
    arrange(FPR, TPR) %>%
    ggplot(aes(FPR, TPR)) +
    geom_line()

roc %>%
    filter(!is.na(precision)) %>%
    arrange(TPR, desc(precision)) %>%
    ggplot(aes(TPR, precision)) +
    geom_line()
```

# Global Score Plot

Each bar is the mean of the absolute values of all scores for a given feature.

Since the mean for the scores within a given features is set to 0 during model
fitting, this is equivalent to average absolute deviation (AAD).

```{r, echo = FALSE}
ggplot(global_df, aes(score, reorder(variable, score))) +
    geom_col() +
    xlab("Score") +
    ylab(NULL)
```

# Intercept

Each feature in the EBM model is "centered" at 0 (meaning the mean prediction
for each feature is 0). This means that the intercept is interpreted as the
value the model will take if all features are "average" (which is in general not
0 as it is with other linear models).

```{r, echo = FALSE, results = "asis", fig.width = 8, fig.height = 4}
tibble(x = "intercept", y = mod$intercept) %>%
    ggplot(aes(x, y)) +
    geom_col() +
    xlab(NULL) +
    ylab("score")
```

# Univariate plots

```{r, echo = FALSE, results = "asis", fig.width = 8, fig.height = 4}

get_truncation <- function(s) {
    run_features[[s]][["visualization"]][["truncate"]]
}

get_split_missing <- function(s) {
    run_features[[s]][["visualization"]][["split_missing"]]
}

get_fill_na <- function(s) {
    run_features[[s]][["fill_na"]]
}

get_plot_type <- function(s) {
    run_features[[s]][["visualization"]][["plot_type"]]
}

truncate_maybe <- function(df, name) {
    t <- get_truncation(name)
    lower <- t[["lower"]]
    upper <- t[["upper"]]
    caption <- if (!is.null(lower) && !is.null(upper)) {
                   sprintf("Truncated from %d to %d", lower, upper)
               } else if (!is.null(lower)) {
                   sprintf("Truncated from %d to -Inf", lower)
               } else if (!is.null(upper)) {
                   sprintf("Truncated from -Inf to %d", upper)
               }
    .df <- if (is.null(lower) && is.null(upper)) {
               df
           } else {
               .lower <- if (is.null(lower)) min(df$value) else lower
               .upper <- if (is.null(upper)) max(df$value) else upper
               filter(df, .lower <= value & value <= .upper)
           }
    list(df = .df, lower = lower, upper = upper, caption = caption)
}

null2alt <- function(default, x) {
    if (is.null(x)) default else x
}

null2na <- function(x) {
    null2alt(NA, x)
}

make_integer_plot <- function(df, name, lower = NULL, upper = NULL,
                              ylab = "Score") {
    fill_cols <- c("score", "stdev")
    .lower <- null2alt(min(df$value), lower)
    .upper <- null2alt(max(df$value), upper)
    .join <- tibble(value = .lower:.upper)
    mutate(df, value = ceiling(value)) %>%
        right_join(.join, by = "value") %>%
        arrange(value) %>%
        fill(all_of(fill_cols), .direction = "downup") %>%
        ggplot(aes(value, score)) +
        geom_col() +
        xlab(name) +
        ylab(ylab) +
        geom_errorbar(aes(ymin = score - stdev, ymax = score + stdev))
}

## TODO use inverse logit here?
make_fraction_plot <- function(df) {
    df %>%
        group_by(value) %>%
        summarize(frac = mean(label),
                  stderr = sqrt(frac * (1 - frac) / n())) %>%
        ggplot(aes(value, frac)) +
        geom_point() +
        geom_errorbar(aes(ymin = frac - stderr, ymax = frac + stderr),
                      width = 0.1)
}

make_integer_fraction_plot <- function(df, name, lower = NULL, upper = NULL) {
    .name <- sym(name)
    df %>%
        mutate(value = ceiling({{ .name }})) %>%
        make_fraction_plot() +
        xlab(name) +
        ylab("Frac(TP)") +
        coord_trans(xlim = c(null2na(lower), null2na(upper)))
}

make_continuous_plot <- function(df, name, ylab = "Score") {
    ggplot(df, aes(value, score)) +
        geom_step(aes(y = score + stdev), color = "red") +
        geom_step(aes(y = score - stdev), color = "red") +
        geom_step() +
        xlab(name) +
        ylab(ylab)
}

make_continuous_fraction_plot <- function(df, name) {
    .name <- sym(name)
    df %>%
        ## TODO average the bin ends to make the axis cleaner
        mutate(value = cut({{ .name }}, 20)) %>%
        make_fraction_plot() +
        xlab(name) +
        ylab("Frac(TP)") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
}

make_categorical_plot <- function(df, name, ylab = "Score") {
    df %>% 
        ggplot(aes(factor(value), score)) +
        geom_col() +
        geom_errorbar(aes(ymin = score - stdev, ymax = score + stdev), width = 0.1) +
        xlab(name) +
        ylab(ylab)
}

make_categorical_fraction_plot <- function(df, name) {
    .name <- sym(name)
    df %>%
        mutate(value = factor({{ .name }})) %>%
        make_fraction_plot() +
        xlab(name) +
        ylab("Frac(TP)")
}

standardize_y_axes <- function(ps) {
    lims <- map(ps, ~ layer_scales(.x)[["y"]]$get_limits()) %>%
        do.call(cbind, .)
    new <- c(min(lims[1, ]), max(lims[2, ]))
    map(ps, ~ .x + ylim(new))
}

make_split_plot <- function(df, name, bound, fun) {
    missing_val <- get_fill_na(name)
    missing <- filter(df, value == missing_val) %>%
        mutate(value = "Missing")
    nonmissing <- filter(df, value != missing_val) %>%
        mutate(value = if_else(value < bound, bound, value))
    bar <- ggplot(missing, aes(factor(value), score)) +
        geom_col() +
        geom_errorbar(aes(ymax = score + stdev,
                          ymin = score - stdev),
                      width = 0.2) +
        xlab(NULL)
    step <- fun(nonmissing, NULL) +
        ylab(NULL) +
        theme(axis.text.y = element_blank(),
              axis.ticks.y = element_blank())
    list(bar, step) %>%
        standardize_y_axes() %>%
        ggarrange(plotlist = ., ncol = 2, widths = c(1, 5)) %>%
        annotate_figure(bottom = text_grob(name))
}

make_split_fraction_plot <- function(df, name, bound, fun) {
    .name <- sym(name)
    missing_val <- get_fill_na(name)
    missing <- filter(df, {{ .name }} == missing_val) %>%
        mutate({{ .name }} := "Missing")
    nonmissing <- filter(df, {{ .name }} != missing_val)
    bar <- make_categorical_fraction_plot(missing, name) +
        xlab(NULL)
    step <- fun(nonmissing, name) +
        xlab(NULL) +
        ylab(NULL) +
        theme(axis.text.y = element_blank(),
              axis.ticks.y = element_blank())
    list(bar, step) %>%
        standardize_y_axes() %>%
        ggarrange(plotlist = ., ncol = 2, widths = c(1, 5), align = "h") %>%
        annotate_figure(bottom = text_grob(name))
}

wrap_split_maybe <- function(name, split_f, f) {
    s <- get_split_missing(name)
    if (is.null(s)) f else partial(split_f, fun = f, bound = s)
}

print_uv_plot <- function(vartype, df, name) {
    r <- if (vartype == "continuous") {
             tr <- truncate_maybe(df, name)
             t <- get_plot_type(name)
             fs <- if (t == "step") {
                       list(
                           make_continuous_plot,
                           make_continuous_fraction_plot
                       )
                   } else if (t == "bar") {
                       list(
                           partial(
                               make_integer_plot,
                               lower = tr[["lower"]],
                               upper = tr[["upper"]]
                           ),
                           partial(
                               make_integer_fraction_plot,
                               lower = tr[["lower"]],
                               upper = tr[["upper"]]
                           )
                       )
                   } else {
                       stop(sprintf("wrong type, dummy; got %s", t))
                   }
             ## TODO only continuous plots can be split (for now)
             list(
                 feat_f = wrap_split_maybe(name, make_split_plot, fs[[1]]),
                 frac_f = wrap_split_maybe(name, make_split_fraction_plot, fs[[2]]),
                 df = tr[["df"]],
                 caption = tr[["caption"]]
             )
         } else if (vartype == "categorical") {
             list(
                 feat_f = make_categorical_plot,
                 frac_f = make_categorical_fraction_plot,
                 df = df,
                 caption = NULL
             )
         } else {
             stop(sprintf("wrong plot type, dummy; got %s", vartype))
         }
    p0 <- r$feat_f(r$df, name)
    p1 <- r$frac_f(alltrain, name)
    cat(sprintf("## %s\n", name))
    print(p0)
    cat("\n\n")
    print(p1)
    cat("\n\n")
    if (!is.null(r$caption)) {
        cat(sprintf("%s\n\n", r$caption))
    }
}

walk(univariate, ~ print_uv_plot(.x$meta$vartype, .x$df, .x$meta$name))

```

# Bivariate plots

```{r, echo = FALSE, results = "asis", fig.width = 8, fig.height = 4}

cont_cont_plot <- function(df, yvar, left_name, right_name) {
    # poor-mans 2d step heatmap plot thing
    df %>%
        group_by(right_value) %>%
        mutate(left_upper = lead(left_value)) %>%
        ungroup() %>%
        group_by(left_value) %>%
        mutate(right_upper = lead(right_value)) %>%
        ungroup() %>%
        filter(!is.na(left_upper)) %>%
        filter(!is.na(right_upper)) %>%
        ggplot() +
        geom_rect(aes_string(xmin = "left_value",
                             xmax = "left_upper",
                             ymin = "right_value",
                             ymax = "right_upper",
                             fill = yvar)) +
        xlab(left_name) +
        ylab(right_name)
}

print_cont_cont_plot <- function(df, left_name, right_name) {
    x_tr <- df %>%
        rename(value = left_value) %>%
        truncate_maybe(left_name)
    y_tr <- df %>%
        rename(value = right_value) %>%
        truncate_maybe(right_name)
    lims <- coord_trans(xlim = c(x_tr$lower, x_tr$upper),
                        ylim = c(y_tr$lower, y_tr$upper))
    if (!is.null(x_tr$caption)) {
        cat(sprintf("%s: %s\n\n", left_name, x_tr$caption))
    }
    if (!is.null(y_tr$caption)) {
        cat(sprintf("%s: %s\n\n", right_name, y_tr$caption))
    }
    cat("### Scores\n\n")
    p0 <- cont_cont_plot(df, "score", left_name, right_name) +
        scale_fill_gradient2(midpoint = 0) +
        lims
    print(p0)
    cat("\n\n")
    cat("### Stdevs\n\n")
    p1 <- cont_cont_plot(df, "stdev", left_name, right_name) +
        scale_fill_gradient() +
        lims
    print(p1)
}

print_cont_cat_plot_inner <- function(df, cat_name, cont_name) {
    cat_val <- as.character(df$c[[1]])
    tr <- truncate_maybe(df, cont_name)
    t <- get_plot_type(cont_name)
    f <- if (t == "step") {
             partial(make_continuous_plot)
         } else if (t == "bar") {
             partial(make_integer_plot,
                     lower = tr[["lower"]],
                     upper = tr[["upper"]]
                     )
         } else {
             stop(sprintf("wrong plot type: got %s", t))
         }
    p <- wrap_split_maybe(cont_name, make_split_plot, f)(tr[["df"]], cont_name)
    cat(sprintf("### %s = %s\n\n", cat_name, cat_val))
    cat(sprintf("%s\n\n", tr[["caption"]]))
    print(p)
    cat("\n\n")
}

print_cont_cat_plot <- function(df, cat, cont, cat_name, cont_name) {
    ## this 'all_of' thing is needed to silence a weird warning about
    ## using vectors to select things (I disagree with it, but whatever)
    .df <- rename(df, value = all_of(cont), c = all_of(cat)) %>%
        mutate(lower = score - stdev,
               upper = score + stdev) %>%
        group_split(c) %>%
        walk(~ print_cont_cat_plot_inner(.x, cat_name, cont_name))
}

print_cat_cat_plot <- function(df, left_name, right_name) {
    p <- ggplot(df, aes(factor(left_value),
                         score,
                         fill = factor(right_value)
                         )) +
        geom_col(position = "dodge") +
        geom_errorbar(aes(ymin = score - stdev,
                          ymax = score + stdev),
                      width = 0.1,
                      position = position_dodge(0.9)) +
        xlab(left_name) +
        scale_fill_discrete(name = right_name)
    print(p)
}

print_bv_plot_inner <- function(L, R, df) {
    if (L$type == "continuous" && R$type == "continuous") {
        print_cont_cont_plot(df, L$name, R$name)
    } else if (L$type == "categorical" && R$type == "continuous") {
        print_cont_cat_plot(df, "left_value", "right_value", L$name, R$name)
    } else if (L$type == "continuous" && R$type == "categorical") {
        print_cont_cat_plot(df, "right_value", "left_value", R$name, L$name)
    } else if (L$type == "categorical" && R$type == "categorical") {
        print_cat_cat_plot(df, L$name, R$name)
    } else {
        sprintf("Types are wrong, dummy: %s and/or %s", L$type, R$type)
    }
}

print_bv_plot <- function(L, R, df) {
    cat(sprintf("## %s x %s\n\n", L$name, R$name))

    print_bv_plot_inner(L, R, df)

    cat("\n\n")
}

if (length(bivariate) == 0) {
    cat("None\n\n")
} else {
    walk(bivariate, ~ print_bv_plot(.x$left, .x$right, .x$df))
}

```

# Univariate Error profiles

```{r, echo = FALSE, results = "asis", fig.width = 8, fig.height = 4}

logErr <- function(p) {
    -log10(1 - p)
}

make_perf_df <- function(df) {
    df %>%
        group_by(value) %>%
        summarize(precision = sum(label & pred) / sum(pred),
                  recall = sum(label & pred) / sum(label),
                  f1 = 2 * (precision * recall) / (precision + recall)) %>%
        pivot_longer(cols = c(precision, recall, f1),
                     names_to = "metric",
                     values_to = "mvalue") %>%
        mutate(mvalue = logErr(mvalue))
        ## ggplot(aes(value, mvalue, color = metric)) +
        ## labs(x = "Feature Value",
        ##      y = "-log10(metric)")
}

make_perf_plot <- function(df) {
    df %>%
        group_by(value) %>%
        summarize(precision = sum(label & pred) / sum(pred),
                  recall = sum(label & pred) / sum(label),
                  f1 = 2 * (precision * recall) / (precision + recall)) %>%
        pivot_longer(cols = c(precision, recall, f1),
                     names_to = "metric",
                     values_to = "mvalue") %>%
        mutate(mvalue = logErr(mvalue)) %>%
        ggplot(aes(value, mvalue, color = metric)) +
        xlab(NULL) +
        ylab("-log10(metric)")
}

make_integer_perf_plot <- function(df, name, lower = NULL, upper = NULL) {
    .n <- sym(name)
    df %>%
        mutate(value = ceiling({{ .n }})) %>%
        make_perf_plot() +
        geom_point() +
        geom_line() +
        coord_trans(xlim = c(null2na(lower), null2na(upper)))
}

make_continuous_perf_plot <- function(df, name) {
    .n <- sym(name)
    df %>%
        mutate(value = cut({{ .n }}, 20)) %>%
        make_perf_plot() +
        geom_point() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
}

make_categorical_perf_plot <- function(df, name) {
    .n <- sym(name)
    df %>%
        mutate(value = factor({{ .n }})) %>%
        make_perf_plot() +
        geom_col(position = "dodge", aes(fill = metric))
}

make_perf_split_plot <- function(df, name, bound, fun) {
    .n <- sym(name)
    missing_val <- get_fill_na(name)
    missing <- filter(df, {{ .n }} == missing_val) %>%
        mutate({{ .n }} := "Missing")
    nonmissing <- filter(df, {{ .n }} != missing_val) %>%
        mutate({{ .n }} := if_else({{ .n }} < bound, bound, {{ .n }}))
    bar <- make_categorical_perf_plot(missing, name) +
        xlab(NULL)
    step <- fun(nonmissing, name) +
        ylab(NULL) +
        xlab(NULL) +
        theme(axis.text.y = element_blank(),
              axis.ticks.y = element_blank())
    list(bar, step) %>%
        standardize_y_axes() %>%
        ggarrange(plotlist = ., ncol = 2, widths = c(1, 5),
                  common.legend = TRUE, legend = "right", align = "h") %>%
        annotate_figure(bottom = text_grob(name))
}

print_perf_profile_plot <- function(vartype, name) {
    r <- if (vartype == "continuous") {
             tr <- get_truncation(name)
             t <- get_plot_type(name)
             f <- if (t == "step") {
                      make_continuous_perf_plot
                  } else if (t == "bar") {
                      partial(
                          make_integer_perf_plot,
                          lower = tr[["lower"]],
                          upper = tr[["upper"]]
                      )
                  } else {
                      stop(sprintf("wrong type, dummy; got %s", t))
                  }
             list(
                 feat_f = wrap_split_maybe(name, make_perf_split_plot, f),
                 caption = tr[["caption"]]
             )
         } else if (vartype == "categorical") {
             list(
                 feat_f = make_categorical_perf_plot,
                 caption = NULL
             )
         } else {
             stop(sprintf("wrong plot type, dummy; got %s", vartype))
         }
    cat(sprintf("## %s\n", name))
    print(r$feat_f(alltrain, name))
    cat("\n\n")
    if (!is.null(r$caption)) {
        cat(sprintf("%s\n\n", r$caption))
    }
}

walk(univariate, ~ print_perf_profile_plot(.x$meta$vartype, .x$meta$name))

```
